Making Deep Q-learning Methods Robust to Time Discretization

### 2019-01-18 paper/icml_drau/advup.tex
.

Eq.~\eqref{eq:max_A}
is directly verified by $A_{\psi}$, owing to the reparametrization
$A_\psi(s, a) = \bar{A}_\psi(s, a) - \bar{A}_\psi(s, \pi(s))$, described
in~\ref{subsec:reparam}.  To approximately verify~\eqref{eq:bellman_A}, the
corresponding squared residual is minimized by an approximate gradient descent.
The update equations when learning from a transition $(s, a, r, s')$, either from
an exploratory trajectory or from a replay buffer~\cite{dqn} are
\begin{align}
	\delta Q_\deltat &\leftarrow A(s, a) \deltat - \left(r \deltat + \gamma^{\deltat} V(s') - V(s)\right)\\
	\theta_\deltat &\leftarrow \theta_\deltat + \eta^V_\deltat \partial_{\theta} V(s) \frac{\delta Q_\deltat}{\deltat}\\
	\psi_\deltat &\leftarrow \psi_\deltat + \eta^A_\deltat \partial_{\psi} A(s, a) \frac{\delta Q_\deltat}{\deltat}.
	\label{eq:approx_bellman_A}
\end{align} 
where $\eta$'s are learning rates.
Appropriate scalings for the learning rates $\eta^V_\deltat$ and $\eta^A_\deltat$ in term of $\deltat$ to obtain a well defined continuous limit are derived in the following section. \TODO{Change notations for the learning rates in the algorithm}
% As we will see in the following section,
% to obtain a well defined continuous time limit, the learning rates can be rewritten as
% $\eta^i_\deltat = \alpha^i \deltat$, where $\alpha^1$ and $\alpha^2$ are independant
% of $\deltat$.
\subsection{Learning rates scalings}
\label{subsec:lr}
\input{pendulum_fig.tex}
For the discrete algorithm to admit a continuous time limit, the trajectories
of parameters must converge to well defined trajectories as $\deltat$ goes to
$0$.  Intuitively, this means that optimization steps should not be too large,
in which case parameter trajectories could diverge in a single gradient step,
or too small, in which case the trajectories would converge to single points as
$\deltat$ goes to $0$.  This in turn imposes conditions on the scalings of the
learning rates of algorithms robust to changes of the tine discretization.

@@

For parameter updates to follow a SDE, one must
have $\eta^V_\deltat = \alpha^V \deltat$ and $\eta^A_\deltat = \alpha^A
\deltat$. More precisely
\begin{theorem}
	\TODO{We need a bit more details: we probably want to assume that states and actions are
	drawn from a single fixed trajectory, and that we are learning for a fixed, constant policy.}
	Let $\eta^V_\deltat = \alpha^V \deltat^\beta$ (resp. $\eta^A \deltat^\beta$) be the learning rate
	of $V$ (resp. $A$), when training on a single exploratory trajectory. Then
	\begin{itemize}
		\item If $\beta = 1$ the discrete parameter trajectories converge to continuous parameter
			trajectories as $\deltat$ goes to $0$.
		\item If $\beta < 1$, parameters grow arbitrarily large in a single timestep as $\deltat$
			goes to $0$.
		\item If $\beta > 1$, parameter trajectories shrink to a single point as 
			$\deltat$ goes to $0$.
	\end{itemize}
\end{theorem}

% With this choice of learning rates, the limit SDEs are
% \begin{align}
% 	F &= A(s_t, a_t) - r_t - \log(\gamma) V(s_t) - \partial_s V(s_t) f(s_t, a_t)\nonumber\\
% 	G &= - \partial_s V(s_t, a_t) \Sigma(s_t, a_t)\nonumber\\
% 	d\theta_t &= (\alpha^V Fdt  + D \dbt)\partial_{\theta} V(s_t)\nonumber\\
% 	d\psi_t &= (\alpha^A Fdt  + D \dbt)\partial_{\psi} A(s_t, a_t)\nonumber.
% \end{align}
The full algorithm can now be derived.
Pseudocode for the discrete actions algorithm is given in Alg.~\ref{alg:dau}.
Pseudocode for the continuous actions algorithm is given in the supplementary
material.
\TODO{Be careful when using RMSprop. When you are facing a stochastic environment
	RMSprop evaluates the second momentum of gradients and rescale gradients by its
	square root. This second momentum will
	be of order $\frac{1}{\deltat}$, due to the stochasticity in the environment. Consequently
	all gradient steps are multiplied by $\sqrt{\deltat}$, and learning rates need only be
multiplied by $\sqrt{\deltat}$ instead of $\deltat$. Taking batches of size proportional to $\frac{1}{\deltat}$
applies a similar fix.}

>>>>

Eq.~\eqref{eq:max_A}
is directly verified by $A_{\psi}$, owing to the reparametrization
$A_\psi(s, a) = \bar{A}_\psi(s, a) - \bar{A}_\psi(s, \pi(s))$, described
in~\ref{subsec:reparam}.  To approximately verify~\eqref{eq:bellman_A}, the
corresponding squared residual is minimized by an approximate gradient descent.
The update equations when learning from a transition $(s, a, r, s')$, either from
an exploratory trajectory or from a replay buffer~\cite{dqn} are
\begin{align}
	\delta Q_\deltat &\leftarrow A(s, a) \deltat - \left(r \deltat + \gamma^{\deltat} V(s') - V(s)\right)\\
	\theta_\deltat &\leftarrow \theta_\deltat + \eta^V_\deltat \partial_{\theta} V(s) \frac{\delta Q_\deltat}{\deltat}\\
	\psi_\deltat &\leftarrow \psi_\deltat + \eta^A_\deltat \partial_{\psi} A(s, a) \frac{\delta Q_\deltat}{\deltat}.
	\label{eq:approx_bellman_A}
\end{align} 
where $\eta$'s are learning rates.
Appropriate scalings for the learning rates $\eta^V_\deltat$ and $\eta^A_\deltat$ in term of $\deltat$ to obtain a well defined continuous limit are derived in the following section.
% As we will see in the following section,
% to obtain a well defined continuous time limit, the learning rates can be rewritten as
% $\eta^i_\deltat = \alpha^i \deltat$, where $\alpha^1$ and $\alpha^2$ are independant
% of $\deltat$.
\subsection{Learning rates scalings}
\label{subsec:lr}
\input{pendulum_fig.tex}
For the discrete algorithm to admit a continuous time limit, the trajectories
of parameters must converge to well defined trajectories as $\deltat$ goes to
$0$.  Intuitively, this means that optimization steps should not be too large,
in which case parameter trajectories could diverge in a single gradient step,
or too small, in which case the trajectories would converge to single points as
$\deltat$ goes to $0$.  This in turn imposes conditions on the scalings of the
learning rates of algorithms robust to changes of the tine discretization.

@@

For parameter updates to follow a SDE, one must
have $\eta^V_\deltat = \alpha^V \deltat$ and $\eta^A_\deltat = \alpha^A
\deltat$. More precisely
\begin{theorem}
	\TODO{We need a bit more details: we probably want to assume that states and actions are
	drawn from a single fixed trajectory, and that we are learning for a fixed, constant policy.}
	Let $\eta^V_\deltat = \alpha^V \deltat^\beta$ (resp. $\eta^A \deltat^\beta$) be the learning rate
	of $V$ (resp. $A$), when training on a single exploratory trajectory. Then
	\begin{itemize}
		\item If $\beta = 1$ the discrete parameter trajectories converge to continuous parameter
			trajectories as $\deltat$ goes to $0$.
		\item If $\beta < 1$, parameters grow arbitrarily large in a single timestep as $\deltat$
			goes to $0$.
		\item If $\beta > 1$, parameter trajectories shrink to a single point as 
			$\deltat$ goes to $0$.
	\end{itemize}
	\label{th:cont-params}
\end{theorem}

% With this choice of learning rates, the limit SDEs are
% \begin{align}
% 	F &= A(s_t, a_t) - r_t - \log(\gamma) V(s_t) - \partial_s V(s_t) f(s_t, a_t)\nonumber\\
% 	G &= - \partial_s V(s_t, a_t) \Sigma(s_t, a_t)\nonumber\\
% 	d\theta_t &= (\alpha^V Fdt  + D \dbt)\partial_{\theta} V(s_t)\nonumber\\
% 	d\psi_t &= (\alpha^A Fdt  + D \dbt)\partial_{\psi} A(s_t, a_t)\nonumber.
% \end{align}
The full algorithm can now be derived.
Pseudocode for the discrete actions algorithm is given in Alg.~\ref{alg:dau}.
Pseudocode for the continuous actions algorithm is given in the supplementary
material.

### 2019-01-18 paper/icml_drau/experiments.tex
Commiting figures.
.

%! TEX root = icml_drau.tex
\section{Discussions}
\label{sec:discussions}

TODO

>>>>

%! TEX root = icml_drau.tex
\section{Experiments}
\label{sec:exp}
\input{lc_fig.tex}
The experiment set that we provide here is specifically aimed at showing that
the proposed method can operate efficiently on a wide range of time
discretization, without specific tuning, while standard deep Q-learning
approaches can't.  Our method, which is referred to as \emph{Deep Advantage
Updating}, is compared to DDPG in continuous action settings and to DQN in
discrete action settings.

In all experimental settings, we use the learning algorithm described in
alg.~\cite{alg:dau} and \TODO{supplementary, algo 1}. The variant of DDPG and DQN
used are described in the supplementary material, as well as all the hyperparameters
used.

Let us stress that quantities used are rescaled to make comparison possible. For example,
return results are given in term of macroscopic reward\footnote{This mostly amounts to scaling rewards
by a factor $\deltat$ when this scaling is not naturally done in the environment. Environment specific
details are given in the supplementary material.}, and, most notably, time elapsed is always given in
normalized epochs, i.e. the true number of epochs multiplied by the discretization timestep.

To provide qualitative results, and check that robustness to time
discretization is provided both in term of raw return results, but also in term
of convergence of approximate value function and policies to the optimal
value function and policy, we first provide results on the simple pendulum environment
provided in the openai gym classic control suite. Qualitative results are presented in
fig.~\ref{fig:full_results}.
% \begin{figure}[h]
%   \centering
%   \includegraphics[width=\columnwidth]{figs_data/cartpole_lc.png}
%   \caption{Cartpole}
%   \label{fig:cartpole-lc}
% \end{figure}

% \begin{figure}[h]
%   \centering
%   \includegraphics[width=\columnwidth]{figs_data/ant_lc.png}
%   \caption{Ant}
%   \label{fig:ant-lc}
% \end{figure}
% 
% \begin{figure}[h]
%   \centering
%   \includegraphics[width=\columnwidth]{figs_data/pendulum_lc.png}
%   \caption{Pendulum}
%   \label{fig:pendulum-lc}
% \end{figure}

### 2019-01-18 paper/icml_drau/framework.tex
.

% TODO: For now, differential equation is the default, SDE is a footnote.
        Let ${\cal S} = \mathbb{R}^d$ be a set of states, and ${\cal A}$ be a set of
	actions. Consider the continuous \emph{Markov Decision Process} (MDP) defined by the \emph{Differential Equation}: %\TODO{Explain the intuition of Brownian in SDE}:
        \begin{equation}
		\label{eq:diffusion}
		ds = F(s, a)dt. % + \Sigma(s, a)\dbt.
              \end{equation}
        For any $\deltat$, the discrete MDP ${\cal
	M}_\deltat = \langle {\cal S}, {\cal A}, T_{\deltat}, r_\deltat,
      \gamma_\deltat\rangle$ is a discretization of the continuous MDP with
      \emph{time discretization} $\deltat$. The transition probability
      $T_\deltat(s, a, s')$ is a dirac on $s_\deltat$, where $s_\deltat$ is the
      state obtained
      when starting at $s_0 = s$ and maintaining $a_t=a$ constant for a time
      $\deltat$.
      This corresponds to an agent evolving in the continuous
      environment described by Eq. \eqref{eq:diffusion}, but who is
      only making observations and choosing actions every $\deltat$. We
      say the the MDP ${\cal M}_\deltat$ is near-continuous.\footnote{
	      We believe the results presented here to be true in a more general framework, where states
	      follow a \emph{Stochastic Differential Equation} $ds = F(s, a)dt.  + \Sigma(s, a)\dbt$, with
	      $B_t$ a multidimensional brownian motion and $\Sigma$ its associated covariance matrix. However,
      formal treatment of SDEs is generally tedious and out of the scope of this paper.}
      Informally $F: {\cal S}\times{\cal A}\rightarrow {\cal S}$ describes
	the dynamic of the environment. The agent interacts with the environment through a deterministic policy function
	$\pi \colon {\cal S} \rightarrow {\cal A}$. Determinism of the policy is assumed to prevent problematic definitions
	of the value function in continuous time. This is only a mild requirement, as we introduce an off-policy algorithm,
	and only impose determinism of the exploitation policy.
	% This framework covers classical control theory, as well as many usual simulated environments. TODO: not convinced that this is relevant here.

>>>>

% TODO: For now, differential equation is the default, SDE is a footnote.
        Let ${\cal S} = \mathbb{R}^d$ be a set of states, and ${\cal A}$ be a set of
	actions. Consider the continuous \emph{Markov Decision Process} (MDP) defined by the \emph{Differential Equation}: %\TODO{Explain the intuition of Brownian in SDE}:
        \begin{equation}
		\label{eq:diffusion}
		ds = F(s, a)dt. % + \Sigma(s, a)\dbt.
              \end{equation}
        For any $\deltat$, the discrete MDP ${\cal
	M}_\deltat = \langle {\cal S}, {\cal A}, T_{\deltat}, r_\deltat,
      \gamma_\deltat\rangle$ is a discretization of the continuous MDP with
      \emph{time discretization} $\deltat$. The transition probability
      $T_\deltat(s, a, s')$ is a dirac on $s_\deltat$, where $s_\deltat$ is the
      state obtained
      when starting at $s_0 = s$ and maintaining $a_t=a$ constant for a time
      $\deltat$.
      This corresponds to an agent evolving in the continuous
      environment described by Eq. \eqref{eq:diffusion}, but who is
      only making observations and choosing actions every $\deltat$. We
      say the the MDP ${\cal M}_\deltat$ is near-continuous.\footnote{
	      We believe the results presented here to be true in a more general framework, where states
	      follow a \emph{Stochastic Differential Equation} \begin{equation}
		      ds = F(s, a)dt.  + \Sigma(s, a)\dbt
		      \label{eq:sde}
      \end{equation}, with
	      $B_t$ a multidimensional brownian motion and $\Sigma$ its associated covariance matrix. However,
      formal treatment of SDEs is generally tedious and out of the scope of this paper.}
      Informally $F: {\cal S}\times{\cal A}\rightarrow {\cal S}$ describes
	the dynamic of the environment. The agent interacts with the environment through a deterministic policy function
	$\pi \colon {\cal S} \rightarrow {\cal A}$. Determinism of the policy is assumed to prevent problematic definitions
	of the value function in continuous time. This is only a mild requirement, as we introduce an off-policy algorithm,
	and only impose determinism of the exploitation policy.
	% This framework covers classical control theory, as well as many usual simulated environments. TODO: not convinced that this is relevant here.
